Firstly, I would express my sincere gratitude to Prof. Yang Liu and Prof Alwen Fernanto Tiu, for their constant encouragement and guidance during my Ph.D years. Prof. Liu has always been supportive and responsible towards my research and personal life both for my PhD years and the year when I was working as a research assistant. He gave me great freedom to explore research topics and provided as many as possible resources to expand my horizon as a researcher. He is also generous to help me rebuild confidence and overcome difficulties when I was confused about my research. Prof. Tiu's rigorous attitude towards research and the critical thinking have left a deep impression on me and inspired me greatly in tackling intractable research problems. I really appreciate their supervision and have learned much from them.

Secondly, I would also like to thank my collaborators for their support and various guidance. They have been setting good examples for me. I am impressed by the humility and high morality of Prof. Zhiwu Xu, the rigorous writing style of Prof. Bihuan Chen, the problem discovering capability of Prof. Yinxing Xue, the open-minded spirit of Mr. Yuekang Li, and the passionate spirit of Dr. Xiaofei Xie. I also feel lucky to be able to work with other researchers like Prof. Guozhu Meng, Prof. Lei Ma, Dr. Haijun Wang, Mr. Cen Zhang, Mr. Xiuheng Wu.

Thirdly, I would like to express my thanks to Prof. Jun Sun, Prof Shang-Wei Lin, Prof Jianjun Zhao, and Prof Jianqi Shi. They have helped me go through the difficult days and provided me valuable suggestions. In addition, I would like to thank so much to all the friends in Cybersecurity Lab, especially Dr. Yuan Zhou, Dr. Ting Su, Dr. Hao Xiao, Dr. Dongxia Wang, Dr. Junjie Wang, Ms. Xiaoning Du, Mr. Zhengzi Xu, Mr. Ruitao Feng, and many other visiting scholars. Pursuing a Ph.D degree has never been easy, but I have been really enjoying these days in Cybersecurity lab.

Finally, I want to express my utmost gratitude to my family. Having been away from home for studies, I am heavily indebted to my parents and relatives for their unconditional support, understanding and trust over the past years. Their meticulously care and guidance will continue as one of my biggest sources of inspiration and their company are precious treasures in my life.